### Introduction to Stringr

```{r include = FALSE , message=FALSE,echo=FALSE}
library(tidyverse)
```

### tidyverse tweets
tidyweb is a part of the tidyverse, an ecosystem of packages designed with common APIs and a shared philosophy. Learn more at tidyverse.org.


You can get the length of the string with str_length():


```{r}
str_length("abc")
#> [1] 3
#> 
```

This is now equivalent to the base R function <tt>nchar()</tt>. Previously it was needed to work around issues with nchar() such as the fact that it returned 2 for nchar(NA). This has been fixed as of R 3.3.0, so it is no longer so important.

You can access individual character using sub_str(). It takes three arguments: a character vector, a starting position and an end position. Either position can either be a positive integer, which counts from the length, or a negative integer which counts from the right. The positions are inclusive, and if longer than the string, will be silently truncated.

```{r}
x <- c("abcdef", "ghifjk")

# The 3rd letter
str_sub(x, 3, 3)
#> [1] "c" "i"

```

# The 2nd to 2nd-to-last character
str_sub(x, 2, -2)
#> [1] "bcde" "hifj"


You can also use str_sub() to modify strings:

str_sub(x, 3, 3) <- "X"
x
#> [1] "abXdef" "ghXfjk"


To duplicate individual strings, you can use str_dup():

str_dup(x, c(2, 3))
#> [1] "abXdefabXdef"   "ghXfjkghXfjkghXfjk"
Whitespace

Three functions add, remove, or modify whitespace:


str_pad() pads a string to a fixed length by adding extra whitespace on the left, right, or both sides.

x <- c("abc", "defghi")
str_pad(x, 10)
#> [1] "   abc" "defghi"
str_pad(x, 10, "both")
#> [1] "   abc" "  defghi  "


(You can pad with other characters by using the pad argument.)

str_pad() will never make a string shorter:

str_pad(x, 4)
#> [1] " abc"   "defghi"


So if you want to ensure that all strings are the same length (often useful for print methods), combine str_pad() and str_trunc():



x <- c("Short", "This is a long string")

x %>% 
  str_trunc(10) %>% 
  str_pad(10, "right")
#> [1] "Short " "This is..."


The opposite of str_pad() is str_trim(), which removes leading and trailing whitespace:

x <- c("  a   ", "b   ",  "   c")
str_trim(x)
#> [1] "a" "b" "c"
str_trim(x, "left")
#> [1] "a   " "b   " "c"
You can use str_wrap() to modify existing whitespace in order to wrap a paragraph of text so that the length of each line as a similar as possible.



jabberwocky <- str_c(
  "`Twas brillig, and the slithy toves ",
  "did gyre and gimble in the wabe: ",
  "All mimsy were the borogoves, ",
  "and the mome raths outgrabe. "
)

cat(str_wrap(jabberwocky, width = 40))
#> `Twas brillig, and the slithy toves did
#> gyre and gimble in the wabe: All mimsy
#> were the borogoves, and the mome raths
#> outgrabe.
Locale sensitive

A handful of stringr are functions are locale-sensitive: they will perform differently in different regions of the world. These functions case transformation functions:



x <- "I like horses."
str_to_upper(x)
#> [1] "I LIKE HORSES."
str_to_title(x)
#> [1] "I Like Horses."

str_to_lower(x)
#> [1] "i like horses."
# Turkish has two sorts of i: with and without the dot
str_to_lower(x, "tr")
#> [1] "ı like horses."


And string ordering and sorting:
```{r}
x <- c("y", "i", "k")
str_order(x)
#> [1] 2 3 1

str_sort(x)
#> [1] "i" "k" "y"
# In Lithuanian, y comes between i and k
str_sort(x, locale = "lt")
#> [1] "i" "y" "k"
```

The locale always defaults to English to ensure that the default behaviour is identically across systems. Locales always include a two letter ISO-639-1 language code (like “en” for English or “zh” for Chinese), and optionally a ISO-3166 country code (like “en_UK” vs “en_US”). You can see a complete list of available locales by running stringi::stri_locale_list().
