---
title: "UL Pavilion "
subtitle: "Limerick R Workshop Venue"
author: "R Workshop"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(prettydoc)
require(leaflet)
```

### Venue 
* This R Workshop will take place on the UL campus at the UL Pavilion on the Clare campus side of the University of Limerick.
* Please take careful note of the Eircode and plan your journey accordingly.

```{r echo=FALSE, warning=FALSE,message=FALSE}

ULlocations <- data.frame(
  Name = c("UL Pavilion","UL Stables Bus stop","Feeney Way Bus stop","EV Charging Point"),
  long =  c( -8.569646785395351,-8.570751855859147,-8.575987527808632, -8.57382030249169),
  lat =  c(52.67923672830164,52.67280316309208,52.67166466816456,52.67511909550905),
  Type = c("Venue","Bus Stop","Bus Stop","Charging Point"),
  Colour = c("red","orange","orange","green")
)



icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = ULlocations$Colour
)

leaflet(ULlocations) %>% addTiles() %>%
  addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(Name))
```


### Getting there

* Dublin Coach has regulare Services to the UL Campus. However, the busstop is on the opposite side of the campus, and a 10-15 minute walk is required.

### Parking 

* Parking is free on the UL campus on the weekends at the UL Pavilion.
* The UL Pavilion has ramp access. If you require any assistance, please advise the workshop organizers.

### WIFI

* Wi-Fi, can be accessed through the UL guest Wi-Fi network.
* You may require to login with your Email address.

### Catering

* Teas and coffees are not provided, but you may be purchased at the adjoining  Bar/Restaurant.
* Likewise, Lunch may be purchased on site between 12 noon and 3pm.
* Snacks may be purchased on an ongoing basis throughout the day.


#### Preparation

* Please have R, RStudio and relevant R packages installed.
* If advised, materials for the workshop can be downloaded from a code repository.
* This Repository would contain R code segements and data files. It is best to download this repository in advance of the workshop.
* More precise instructions specific to this workshop will be provided seperately.



### Code of Conduction 
* Please note that the R consortium code of conduct will apply.
* Behavior consistent with the modern professional Workplace is expected.

