Using the quantile function
========================================================
The `quantile` function  produces sample quantiles corresponding to the given probabilities. The smallest observation corresponds to a probability of 0 and the largest to a probability of 1.

### Creating a vector 

First we will create two vectors that we will use later.
The semi-colon operator is used to create a sequence of integers.

```{r}
deciles = 0:10/10
quintiles = 0:5/5
deciles
quintiles

```


```{r}
attach(iris)
quantile(Sepal.Length,0.10)
quantile(Sepal.Length,0.90)
quantile(Sepal.Length,c(0.10,0.90))
quantile(Sepal.Length,deciles)
quantile(Sepal.Length,quintiles)
detach(iris)
```
<!------------------------------------------>

### Tukey Five Number Summary

```{r}
attach(iris)
quantile(Sepal.Length,c(0.25,0.50,0.75))
fivenum(Sepal.Length)
detach(iris)
```
There are several different methods of determining a quantile. You can specify an alternative approach to the default method using the `type=` argument. Often the resulting value will still be the same.
```{r}
attach(iris)
quantile(Sepal.Length, 0.5 , type=1:3)
detach(iris)
```
